### 2023-02-11 article.tex
Add data preservation

@@

Again, useful paradata should be comprehensive, 
documented appropriately and in co-evolvution with the data,
and the data should be friendly to computers \cite{huvila2022improving}.
In this section, these ideal properties of paradata are applied to code.
Code as paradata, however, has multiple uses as well,
which are to reproduce, to replicate and to 
extend a computational experiment \cite{benureau2018re}.
In this section, the recommendations how to make paradata useful
is applied to code and compared to existing best practices in software 
development, for the intended use of the code.

@@

For code to be extensive enough to reproduce an existing experiment,
it should be distributed in standard ways \cite{peng2006reproducible}
(see section \ref{sec:code-availability}
for the general availability of code).
The standard way to publish code is to use a code hosting website.
There are multiple code hosting websites, 
with GitHub being the most popular one \cite{cosentino2017systematic}.
GitHub is a website where users can create dedicated pages (called 'repositories')
to upload code and interact with it.
The use of such code hosting websites
has increased strongly \cite{russell2018large},
accommodates collaboration \cite{perez2016ten}
and improves transparency \cite{gorgolewski2016practical}.
See \cite{cosentino2017systematic} for an extensive overview of
research conducted on GitHub.

@@

In software development, 
changes are common and should be embraced \cite{beck2000extreme}.

@@

@@

Preserving code by uploading it to a code hosting webite
has many advantages.
Code is harder to preserve than an English text
and preserving code is rarely done \cite{barnes2010publish}.
Although code is the primary actor in computational experiments,
there is no incentive to submit code alongside a publication.
Most academic journal do not require authors to submit their code,
nor is the submitted code peer reviewed, 
rOpenSci \cite{ram2018community} being the pleasant exception.

>>>>

Preserving code in a reliable way is rare \cite{barnes2010publish}.
In general, within software development, 
changes at all levels are common
and instead of ignoring that, it should be embraced \cite{beck2000extreme}.
This section discusses the preservation of code for a short, medium
and long term.

\subsection{Code hosting}

To preserve code for a short term, 
a code hosting website is a good first step.
A code hosting website is a website where 
its users can create dedicated pages (called 'repositories')
for a project, upload code and interact with that code.
There are multiple code hosting websites, 
with GitHub being the most popular one \cite{cosentino2017systematic}.
The \LaTeX code of this chapter is hosted on GitHub 
as well, at \url{https://github.com/richelbilderbeek/chapter_paradata}.
The use of code hosting websites
has increased strongly \cite{russell2018large},
accommodates collaboration \cite{perez2016ten}
and improves transparency \cite{gorgolewski2016practical}.
See \cite{cosentino2017systematic} for an extensive overview of
research conducted on GitHub.

However, if the code breaks tomorrow, it will go unnoticed.

\subsection{Continuous integration}

To preserve code for a longer timespan, 
continuous integration allows one to 
verify if code still works and, if not, to be notified on the need for an update.

However, for code to work longer, one can put the code and its dependencies in
a container, that should be best in standing the test of time.

\subsection{Containerization}

To preserve code for the longest time, 
both code and its dependencies are put into a so-called container.

@@

To repeat features of useful paradata: paradata should be comprehensive, 
documented appropriately, documented in co-evolution with the data,
and the data should be friendly to computers \cite{huvila2022improving}.
In this section, these ideal properties of paradata are applied to code.
However, code has multiple uses, which have different requirements
for being usesful paradata.
Code can be used to reproduce, to replicate and to 
extend a computational experiment \cite{benureau2018re},
each of these uses has different requirements for the code to be useful paradata.
In this section, the recommendations how to make paradata useful
is applied to code and compared to existing best practices in software 
development, for the intended use of the code.

@@

Code should be hosted.

For code to be extensive enough to reproduce an existing experiment,
it should be distributed in standard ways \cite{peng2006reproducible}
(see section \ref{sec:code-availability}
for the general availability of code).
The standard way to publish code is to use a code hosting website.

@@

@@

The most reproducible way of submitting the code of an experiment,
is by providing the code with all the (software) dependencies 
it needs installed in a virtual container.

@@
